---
title: "exercise9A"
output: html_document
---


#### Multiple linear regression - additive model
```{r}
loyn <- read.table('Chapter9/Data/loyn.csv', header=T, sep=',')
```

#### Asess assumptions of linearity, normality and homogeneity of variance.
```{r}
library(car)
scatterplotMatrix(~ABUND+AREA+YR.ISOL+DIST+LDIST+GRAZE+ALT, data=loyn, diag="boxplot")
```

*Conclusions - AREA, DIST, LDIST variables non-normal and consequently the relationship show non-linearity. 

#### Scale transformations should be attempted.
```{r}
scatterplotMatrix(~ABUND+log10(AREA)+YR.ISOL+log10(DIST)+
                     log10(LDIST)+GRAZE+ALT, data=loyn, diag="boxplot")
```

* log 10 transfromation appear successful. no evidence of non-normality, non-homogeneity of variacne or non-linearity

#### Assess multicollinearity
```{r}
cor(loyn[,2:7])
```

* Conclusions - With the exception of GRAZE and YR.ISOL, non corrrelations

#### Variance inflation and their inverse(tolerances)
```{r}
vif(lm(ABUND~log10(AREA)+YR.ISOL+log10(DIST)+log10(LDIST) +GRAZE+ALT, data=loyn))
1/vif(lm(ABUND~log10(AREA)+YR.ISOL+log10(DIST)+log10(LDIST) +GRAZE+ALT, data=loyn))
```

* Conclusions - VIF < 5, tolerances > 0.2 means that multicollinearity is unlikely to be a problem.

#### Fit the additive muliple linear model
```{r}
loyn.lm<-lm(ABUND~log10(AREA)+YR.ISOL+log10(DIST)+log10(LDIST)+GRAZE+ALT,
            data=loyn)
```

#### Plotting
```{r}
plot(loyn.lm)
```

* Conclusions - There is no obvious "Wedge" pattern evident in the residual plot. none of the point approach the high Cook's D contours suggesting theat none of the observations are overly influential on the final fitted model.

#### Influence
```{r}
influence.measures(loyn.lm)
```

* Conclusions - leverage(hat) values are greater than 2*p/n = 0.286. none of the Cook's D values are >= 1. Hence the hypothesis tests are likely to be reliable.

#### Summary
```{r}
summary(loyn.lm)
```

* Conclusions - a significant positive partial slope for bird abundance against log10 patch area. The overall model explained 69% of the vairablility

#### Individual partial relationships between the response variable and each of the predictor variable
```{r}
avPlots(loyn.lm, ask=F)
```
